EasyMerge - A New Tool for Code Clones Refactoring

### 2014-03-16 paper/paper.tex
paper update

Clone detection has been an avid research topic in the field of software engineering for decades. Fortunately, several automated techniques for detecting code clones
have already been proposed. However, how to deal with detected clones, e.g. how to distinguish necessary clones from unnecessary ones and how to refactor code to remove
unnecessary clones still remain a big problem in not only commercial but also academic domain. As a result, in this paper, we try to classify code clones and build
a recommendation system called EasyMerge to help developers merge unnecessary clones on top of current state-of-the-art clone detection approach.

More specifically, we pick CloneDigger \cite{bulychev}, an anti-unification duplicate code detection tool as our underlying clone detection approach. CloneDigger is one of the best available 
clone detection tools currently for its overall performance, coverage of multiple clone types, and availability. 
EasyMerge integrates CloneDigger as the pre-processing tool, analyze its output clone pairs, and recommend possible merges which can remove unnecessary clones without changing functionality of code base, creating reference conflicts, nor causing troubles to future code understanding and development.

@@

\subsection{Clone Refactoring}
On the other hand, for clone refactoring, most proposed solutions use some pre-defined metrics as the refactoring guidance to help users determine whether the 
code clones are suitable for refactoring. For example, Balazinska \cite{balazinska} proposed to use 21 metrics to measure the suitability for refactoring. Higo \cite{higo}
developed ARIES tools, using the code detection tool CCFinder and environment analysis tool Gemini to find the clones can be refactored. And Schulze \cite{schulze}
proposed a new metric DIST (distance of the code clones). Unlike them, instead of calculating distance between code fragments and only merging close clones, EasyMerge 
try to fix the differences and merge as many clones as possible without introducing extra complexity.

@@

Unfortunately, code clones may be coupled with surrounding code, its inheritance tree, or other independent classes/code blocks. For instance, code fragments inside clone pairs
may contain function calls or variable references which are not defined inside the fragments. Furthermore, functions and variables of the same names may be in fact different
at different locations of the code base. For example, if we have a code clone fragment as following:

@@

However, this creates two new problems. Firstly, in this example, the scope of b was changed. It is now inside the helper function and no longer available to code after the clone fragment
which could potentially cause reference errors. Secondly, when the length of the clone fragment is too long, we could possibly end up with a helper function of tons of parameters
which in turn make the code after refactoring harder to understand and is against our goal to improve readability. And the most intuitive approach to solve the first problem
is to return b and such variables from the ``Extracted Method" which introduces even more complexity.

As a result, for recently proposed refactoring approach \cite{li} and the tools mentioned at the beginning of this subsection, they count the number of such external references
in a code clone fragment as the key factor of their refactoring guideline metrics. 
And clone fragments of too many such occurrences will not be recommended for refactoring.
Furthermore, in Li's work \cite{li}, after standardization, it can also be used to refactor semantical
similar but syntax-wise different clones by moving different function references to extracted method's parameters.

@@

\section{CloneDigger}
\subsection{Overview}
Techniques for detecting duplicate code can be classified according to several criteria. Code can be viewed as similar based on syntactic criteria or at a semantic level. CloneDigger considers only syntactic similarity. The algorithm is approached based on abstract syntax trees. The algorithm of finding duplicates consists of several phases. In the beginning all statements are partitioned into clusters using anti-unification distance and the code is abstractly viewed as sequence of cluster identifiers. All pairs of identical sequences of cluster IDs, which have similar statements in corresponding positions, are globally checked for similarity using anti-unification distance.

@@

\section{EasyMerge}

>>>>

Clone detection has been an avid research topic in the field of software engineering for decades. Fortunately, several automated techniques for detecting code clones
have already been proposed. However, how to deal with detected clones, e.g. how to distinguish necessary clones from unnecessary ones and how to refactor code to remove
unnecessary clones still remain a big problem in not only commercial but also academic domain. In this paper, we try to classify code clones and build
a recommendation system called EasyMerge to help developers merge unnecessary clones on top of current state-of-the-art clone detection approach.

More specifically, we pick CloneDigger \cite{bulychev}, an anti-unification duplicate code detection tool as our underlying clone detector. CloneDigger is one of the best available 
clone detection tools for its overall performance, coverage of multiple clone types, and availability. 
EasyMerge integrates CloneDigger as the pre-processing tool, analyze its output clone pairs, and recommend possible merges which can remove unnecessary clones without changing functionality of code base, creating reference conflicts, nor causing troubles to future code understanding and reusing.

@@

\subsection{Clone Refactoring}
On the other hand, for clone refactoring, most proposed solutions use pre-defined metrics as the refactoring guidance to help users determine whether the 
code clones are suitable for refactoring. For example, Balazinska \cite{balazinska} proposed to use 21 metrics to measure the suitability for refactoring. Higo \cite{higo}
developed ARIES tools, using the code detection tool CCFinder and environment analysis tool Gemini to find the clones can be refactored. And Schulze \cite{schulze}
proposed a new metric DIST (distance of the code clones). 
%Unlike them, instead of calculating distance between code fragments and only merging close clones, EasyMerge 
%try to fix the differences and merge as many clones as possible without introducing extra complexity.

@@

Unfortunately, code clones may be coupled with surrounding code, its inheritance tree, or other independent classes/code blocks. For instance, code fragments inside clone pairs
may contain function calls or variable references which are not defined inside the fragments. Furthermore, functions and variables of the same names may be in fact different
at different locations of the code base. For example, if we have a code clone fragment of following:

@@

However, this creates two new problems. Firstly, in this example, the scope of b was changed. It is now inside the helper function and no longer available to code after the clone fragment
which could potentially cause reference errors. Secondly, when the length of the clone fragment is too long, we could possibly end up with a helper function of tons of parameters
which in turn make the code after refactoring harder to understand and is against our goal to improve readability. 
For most clone refactoring approaches, this gives bad scores to the evaluation metrics and prevent merging. And a possible alternative approach
is to return b and such variables from the ``Extracted Method" which introduces even more complexity.

For recently proposed refactoring approach \cite{li} and the tools mentioned at the beginning of this subsection, they count the number of such external references
in a code clone fragment as the key factor of their refactoring guideline metrics. 
And clone fragments of too many such occurrences will not be recommended for refactoring.
Furthermore, in Li's work \cite{li}, after standardization, it can also be used to refactor some semantical
similar but syntax-wise different clones by moving different function references to extracted method's parameters.

Additionally, results from clone detectors are usually clone pairs but it's very common 
more than two fragments are clones of each other. Moreover, overlapping clone fragments often
exist in large scale code bases, too. For example, it's not difficult to find a code block of ABCDEF where BC is a clone fragment of one other block and CDE is a clone fragment of another. For both cases, merging one set of clones will in turn break clones in another set. 
And we need to take care of all reference errors caused by code deletion. 
Thus, our clone refactoring tool must handle all these potential issues without
introducing too much computational overhead.

@@

\section{CloneDigger}
CloneDigger \cite{bulychev} was an anti-unification code clone detector introduced by Bulychev and Minea in 2008. It's one of the latest open source clone detection tool.

\subsection{Overview}
Techniques for detecting duplicate code can be classified according to several criteria. Code can be viewed as similar based on syntactic criteria or at a semantic level. CloneDigger considers only syntactic similarity. The algorithm is approached based on abstract syntax trees. The algorithm of finding duplicates consists of several phases. In the beginning all statements are partitioned into clusters using anti-unification distance and the code is abstractly viewed as sequence of cluster identifiers. All pairs of identical sequences of cluster IDs, which have similar statements in corresponding positions, are globally checked for similarity using anti-unification distance.

@@

\section{EasyMerge}
\subsection{User Interface}

\subsection{Preprocessing}
